submission

### 2024-03-09 submission.Rmd
MAde small changes and made it astethically pleasing

```{r,warning=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(MASS)
library(mixtools)
library(pwr)
```

@@

Furthermore, it is worth noting that the sample size between the two groups are different and thus the traditional t-test mechanisms are changed to account for this. However, the techniques used above will be able to handle this internally.

@@

Immediately we can observe that we have significantly less data than for the other exercises, which can indicate that test results can be inconclusive. To verify how many samples we should have to be able to be able to perform such a test, we can use the power of a t-test. To apply this test we have to understand what

# Perform power analysis
n_active <- 4  # Sample size of the active group
n_inactive <- 9  # Sample size of the inactive group

sp <- sqrt(((n_active-1)*sd(diabetic_active$Glu)**2 + (n_inactive - 1)*sd(diabetic_inactive$Glu)**2**2)/(n_active+n_inactive-2))
effect_size <- (mean(diabetic_active$Glu) - mean(diabetic_inactive$Glu))/sp
alpha <- 0.05  # Significance level

# Perform power analysis
pwr.t.test(n = NULL, d = effect_size, sig.level = alpha, power = 0.5,
                              alternative = "two.sided")
```

@@

### 2.1  We first want to study the gene which is the most correlated with the level of cholesterol. Identify this gene and plot `cholesterol` as a function of this gene expression level. Fit a linear model. Test if there is a significant relation between this gene expression and `cholesterol`?

@@

## 2.2  We now wish to perform the previous test for all of the 3116 genes. For each of the genes, fit a linear model that explains `cholesterol` as a function of the gene expression and compute the p-value of the test. AND

## 2.3  Order the p-values and plot the ordered p-values as a function of their rank. On the same plot, display the line $y = x/3116$. Discuss the result.

@@

@@

```
Plotting all values, we can conclude that there are amount of models that have a p-value below 0.05. It is tempting to make the conclusion that all these models are statistically significant since they all have a small p-value. However, in the event that the we only have random data, we still expect 5\% of the models to appear significantly significant. In fact, if we were to only have random data, we expect a uniform distribution between 0 and 1 for the p-values. Visually, looking at the Figure above, we can the p-values perform better that the uniform distribution and therefore appear to contain significant findings.

@@

```
 In the nest question, we are tasked with finding the maximum number of genes such that the expected number of false discoveries is at most 5\%. For this method, we use the Benjamini-Hochberg procedure \cite{FDR}. In short, it states that if we have a set of $\{ H_1, H_2, ... \: , H_m \}$ with associated and ordered p-values $\{ p_1, p_2, ... \: , p_m \}$, then the threshold for the index can be written as:

##2.4

@@

print(index)

@@

```
We can see that $\alpha \frac{i}{m}$ plots out a line, and we choose i such that $P_{(i)}$ always is smaller or equal to that value. Doing so, we get 941 discoveries such that the expected number of false discoveries is no more than 5 \%

##2.5 A stronger guarantee would be that on impose that there shouldn't be more than a 5 \% risk that any discovery is false. For this method, we use the Holm–Bonferroni method. The method is as follows:\\

@@

Using this process, we see that we get 223 discoveries. We also not that smallest p-value in that set is approximately equal to $1.7 \cdot 10^{-5}$.

>>>>

```{r,warning=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(MASS)
library(mixtools)
```

@@

@@

Immediately we can observe that we have significantly less data than for the other exercises, which can indicate that test results can be inconclusive.

@@

### 2.1 Finding the most significant Gene

We first want to study the gene which is the most correlated with the level of cholesterol. Identify this gene and plot `cholesterol` as a function of this gene expression level. Fit a linear model. Test if there is a significant relation between this gene expression and `cholesterol`?

@@

## 2.2 Calculating all p-values

Plotting the p-values of all models, we see that a great number of genes correspond to a p-value of less that 0.05. At a first glance, genes as a whole would appear to contain information that could help us make discoveries.

@@

```{r}
plot(hist(p_list, breaks=20, col = c("orange", rep("lightblue", 100))), main = "Density of p-values", xlab = "p-values")

## 2.3 Ordering p-values

Plotting all values, we can conclude that there are amount of models that have a p-value below 0.05. It is tempting to make the conclusion that all these models are statistically significant since they all have a small p-value. However, in the event that the we only have random data, we still expect 5% of the models to appear significantly significant. In fact, if we were to only have random data, we expect a uniform distribution between 0 and 1 for the p-values. Visually, looking at the Figure below, we can the p-values perform better that the uniform distribution and therefore appear to contain significant findings.

@@

Shuffling the labels, we get the case that the p-values are approximately uniformly distributed and that the p-values follow the line

@@

plot(hist(p_list_shuffeled, breaks=20, col = c("orange", rep("lightblue", 100))), main = "Density of p-values", xlab = "p-values")

## 2.4 Benjamini-Hochberg procedure

In the nest question, we are tasked with finding the maximum number of genes such that the expected number of false discoveries is at most 5%. For this method, we use the Benjamini-Hochberg procedure. In short, it states that if we have a set of $\{ H_1, H_2, ... \: , H_m \}$ with associated and ordered p-values $\{ p_1, p_2, ... \: , p_m \}$, then the threshold for the index can be written as:

@@

@@

We can see that $\alpha \frac{i}{m}$ plots out a line, and we choose i such that $P_{(i)}$ always is smaller or equal to that value. Doing so, we get 941 discoveries such that the expected number of false discoveries is no more than 5 %

## 2.5 Holm–Bonferroni method

A stronger guarantee would be that on impose that there shouldn't be more than a 5 % risk that any discovery is false. For this method, we use the Holm–Bonferroni method. The method is as follows:\\

@@

Using this process, we see that we get 223 discoveries. We also not that biggest p-value in that set is approximately equal to $1.7 \cdot 10^{-5}$.
